
---
title: "Plotting Maps in R"
author: "Simon Adamov"
date: "November 20, 2020"
output: html_document
---
## Setup

This vignette shows some examples on how to plot pretty maps in R with the latest packages.
It should serve as a source of inspiration and sort of cookbook to create simple -> elaborate plots.
The plot below has to be adjusted for each specific input field.
Currently, it functions as a loose collection of open-source and third-party map material 
combined with various raster and shapefiles. 
I would already like to mention that the most beautiful map was completely taken from here: 
https://timogrossenbacher.ch/2019/04/bivariate-maps-with-ggplot2-and-sf/

```{r setup, include=FALSE, echo = FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.height = 10,
  fig.width = 16,
  fig.retina = 3
)

library(here)
library(ncdf4)
library(dplyr)
library(broom)
library(leaflet)
library(raster)
library(OpenStreetMap)
library(ggplot2)
library(ggthemes)
library(htmlwidgets)
library(sf)
library(lintr)
library(cowplot)
library(viridis)
library(readr)
library(magrittr)
library(purrr)
library(stringr)
library(tidyr)
library(conflicted)
library(ggmap)
library(rmapshaper)
library(mapview)
library(rgdal)
# library(languageserver)
# library(httpgd)
# library(jsonlite)

conflict_prefer("filter", "dplyr")
conflict_prefer("select", "dplyr")
```

## Data Import

A typical output format in weather and climate science are netcdf files, hence we import one of this 
type right here, showing the median pollen concentrations modeled by Cosmo-1E.
The code can of course be adjusted to plot any netcdf file. The link below points to the CSCS 
supercomuting center in Locarno (ETH).

Weather models often work with different projections. The code below is using lat/lon values
in standard cartesian coordinates (already unrotated model output).

```{r }
#### SPECIFY YOUR FIELD AND FILE HERE ####
field <- "ALNU"
path_in <- paste0(here(), "/data/")
path_out <- paste0(here(), "/output/")
nc_path <- "/scratch/sadamov/wd/21_alnu_tuning_v3/21022500_c1e_tsa_alnu_tuning/lm_coarse/lfff01000000"
variable  <- "ALNU"
  system(paste0("fxfilter --force -s ", variable, " ", nc_path, " -o ", nc_path, "_filtered"))
  system(paste0("fxconvert --force nc ", nc_path, "_filtered -o ", nc_path, "_filtered.nc"))
nc_path <- paste0(nc_path, "_filtered.nc")
```

```{r data, echo = TRUE}
con <- nc_open(nc_path)
layer <- ncvar_get(con, field)[, , 80]
x <- ncvar_get(con, "lon_1")
y <- ncvar_get(con, "lat_1")
nc_close(con)
```

## Data Preperation

Different packages will require different formats. The most recent development when it comes to 
polygon data in R is the package sf, a highly flexible framework not yet supported by all other packages.
Hence we prepare the data in different shapes and forms. The output of many weather models is a 
gridded field. First, we display such grids. In a second step we extract data from model output 
at specific locations (Swiss municipalities) to create even nicer plots specifically for the Swiss Domain.

In the example below the lat/lon data is first converted into a raster, then aggregated and 
converted into a polygon layer. UPDATE: There are more accurate ways to plot the model gridboxes with
the actual shape. Especially for publications the approach below might not suffice: 
https://stackoverflow.com/questions/43612903/how-to-properly-plot-projected-gridded-data-in-ggplot2
If the reader has time, feel free to implement the methods described in the post above :)

```{r wrangling, echo = TRUE}
# Default raster object
r <- raster(layer,
  ymn = 42.67, ymx = 49.52,
  xmn = 0.16, xmx = 16.75, crs = "+proj=longlat"
)
# A dataframe containing lat/long and values
layer_latlong <- tibble(
  x = c(x),
  y = c(y),
  layer = c(layer)
) %>%
  # lons west of 0 deg. need to be negative, hence we make them negative here
  mutate(x = if_else(x > 180, x - 360, x))
# Convert the banana-shape into an equally spaced grid
layer_raster <- rasterize(cbind(c(x), c(y)),
  r, layer_latlong$layer,
  fun = mean
)
# Plots can become heavy - adjust second agrument accordingly
layer_raster_vcoarse <- aggregate(layer_raster, 20)
layer_raster_coarse <- aggregate(layer_raster, 6)
# Polygons are usually faster to plot than raster images
layer_poly <- rasterToPolygons(layer_raster_coarse)
layer_poly_leaflet <- rasterToPolygons(layer_raster_vcoarse)
# This is the oldschool way before sf was developped
layer_poly@data$id <- seq_len(nrow(layer_poly@data))
poly_fort <- tidy(layer_poly, data = layer_poly@data)
# join data
poly_fort_mer <- merge(poly_fort, layer_poly@data,
  by.x = "id", by.y = "id"
)
# This is the modern way and works well with ggmap
layer_poly_sf <- st_as_sf(layer_poly)
```

## Plots

### Plot Theme

We first define a general map-them (from Timo Grossenbacher) to obtain a similar look and feel for all plots:

```{r }
my_maptheme <- function(...) {
  theme_minimal() +
    theme(
      text = element_text(
        family = "Verdana",
        color = "#22211d"
      ),
      # remove all axes
      axis.line = element_blank(),
      axis.text.x = element_blank(),
      axis.text.y = element_blank(),
      axis.ticks = element_blank(),
      # add a subtle grid
      panel.grid.major = element_line(color = "#dbdbd9", size = 0.2),
      panel.grid.minor = element_blank(),
      # background colors
      plot.background = element_rect(
        fill = "#f5f5f2",
        color = NA
      ),
      panel.background = element_rect(
        fill = "#f5f5f2",
        color = NA
      ),
      legend.background = element_rect(
        fill = "#f5f5f2",
        color = NA
      ),
      # borders and margins
      plot.margin = unit(c(.5, .5, .2, .5), "cm"),
      panel.border = element_blank(),
      panel.spacing = unit(c(-.1, 0.2, .2, 0.2), "cm"),
      # titles
      legend.title = element_text(size = 11),
      legend.text = element_text(
        size = 9, hjust = 0,
        color = "#22211d"
      ),
      plot.title = element_text(
        size = 15, hjust = 0.5,
        color = "#22211d"
      ),
      plot.subtitle = element_text(
        size = 10, hjust = 0.5,
        color = "#22211d",
        margin = margin(
          b = -0.1,
          t = -0.1,
          l = 2,
          unit = "cm"
        ),
        debug = F
      ),
      # captions
      plot.caption = element_text(
        size = 7,
        hjust = .5,
        margin = margin(
          t = 0.2,
          b = 0,
          unit = "cm"
        ),
        color = "#939184"
      ),
      ...
    )
}
```

### Simple and Fast Tile Plot

This is the fastest way to plot a netcdf file. Tiles are being drawn around each coordinate-pair.
We are not looking at base-R plot function here, as ggplot allows for nicer plots and more
customization (imho).

```{r }
ggplot(layer_latlong) +
  geom_tile(aes(x, y, fill = layer, width = 0.018, height = 0.011)) +
  borders("world", xlim = range(layer_latlong$x), ylim = range(layer_latlong$y), colour = "white") +
  scale_fill_viridis_c() +
  theme_bw() +
  coord_fixed(xlim = range(layer_latlong$x), ylim = range(layer_latlong$y)) +
  ggtitle("Simple and Fast Tile Plot") +
  labs(fill = "Legendary") +
  xlab("Longitude [East]") +
  ylab("Latitude [North]") +
  coord_quickmap(xlim = range(layer_latlong$x), ylim = range(layer_latlong$y)) 
```

### Polygon Plot - Open-Source

The following plots are fully open source. They draw polygons on top of a nice map in the background.

```{r efficient, echo = TRUE}
upper_left <- c(51, 0)
lower_right <- c(40, 17.5)
# osm and other map providers available
map <- openmap(upper_left, lower_right, type = "stamen-terrain")
# The raster map is in a mercator projection and must be transformed
map_proj <- openproj(map)
raster_map <- OpenStreetMap::autoplot.OpenStreetMap(map_proj) +
  my_maptheme() +
  geom_polygon(
    data = poly_fort_mer,
    aes(x = long, y = lat, group = group, fill = layer),
    alpha = 0.7,
    size = 0
  ) +
  scale_fill_gradientn("Poaceae\n[m^-3]", colours = rev(heat.colors(200))) +
  labs(
    x = "Lon",
    y = "Lat",
    title = "Grass Pollen in Cosmo-1 Domain",
    subtitle = "Hourly Average Concentration on the 1st of July 2020 at Midnight
    "
  )

raster_map
```

```{r eval = FALSE, include = FALSE}
# Change resolution here
ggsave(paste0(path_out, "raster_map.png"),
  raster_map,
  dpi = "retina"
)
```

```{r}
# Full Domain is c(51, 0), c(41, 18)
# Switzerland is c(48, 6), c(45.5, 10)

# Zoom
upper_left_zoom  <- c(47.5, 8)
lower_right_zoom <- c(46, 10.5)

map_zoom <- openmap(upper_left_zoom, lower_right_zoom, type = "stamen-terrain")
map_proj_zoom <- openproj(map_zoom)
layer_latlong_zoom <- layer_latlong %>%
  filter(
    between(x, upper_left_zoom[2], lower_right_zoom[2]),
    between(y, lower_right_zoom[1], upper_left_zoom[1])
  )

alpha_low <- min(0.2 / (lower_right_zoom[2] - upper_left_zoom[2]), 0.2)
alpha_high <- min(0.7 / (lower_right_zoom[2] - upper_left_zoom[2]), 0.7)

gg_map <- OpenStreetMap::autoplot.OpenStreetMap(map_proj_zoom) +
  my_maptheme() +
  geom_point(aes(x = x, y = y, col = layer, alpha = layer),
    stroke = 1, shape = 15, size = 3.3, data = layer_latlong_zoom
  ) +
  scale_alpha_continuous(range = c(alpha_low, alpha_high)) +
  scale_color_gradientn("Poac", colours = rev(heat.colors(200))) +
  guides(alpha = "none") +
  labs(
    x = "Lon",
    y = "Lat",
    title = "Grass Pollen in Switzerland",
    subtitle = "Hourly Average Concentration on the 1st of July 2020 at Midnight
    "
  )

gg_map
```

```{r eval = FALSE, include = FALSE}
ggsave(paste0(path_out, "gg_map.png"), gg_map, dpi = "retina")
```

### Static - Google-API

The next example requires a Google account to retrieve data from the Google API. 
This package is quite powerful and works well with the modern sf datatype.

```{r ggmap, echo = TRUE}
# ?register_google
# Enter your key here!!!!
register_google(key = "")

centroid <- c(45, 8)
lat_zoom <- c(45, 48)
long_zoom <- c(5, 15)

ggmap_heat <- get_map(c(lon = centroid[2], lat = centroid[1]),
  zoom = 6, maptype = "terrain", color = "color", scale = "auto"
) %>%
  ggmap() +
  geom_sf(aes(fill = layer, alpha = layer),
    data = layer_poly_sf, inherit.aes = FALSE, lwd = 0
  ) +
  my_maptheme() +
  theme(panel.grid.major = element_line(color = "white")) +
  scale_fill_gradientn("Grass\nPollen", colors = rev(heat.colors(200))) +
  scale_alpha_continuous(range = c(0.1, 0.9)) +
  scale_x_continuous(limits = long_zoom, expand = c(0, 0)) +
  scale_y_continuous(limits = lat_zoom, expand = c(0, 0)) +
  guides(alpha = "none") +
  labs(
    x = "Lon",
    y = "Lat",
    title = "Grass Pollen in Switzerland",
    subtitle = "Hourly Average Concentration on the 1st of July 2020 at Midnight
    "
  )

ggmap_heat
```

```{r eval = FALSE, include = FALSE}
ggsave(paste0(path_out, "ggmap_heat.png"), ggmap_heat, dpi = "retina")
```

Changing the color gradient is easy. https://www.datanovia.com/en/blog/top-r-color-palettes-to-know-for-great-data-visualization/

```{r}
centroid <- c(45, 8)
lat_zoom <- c(45, 48)
long_zoom <- c(0, 15)

ggmap_viridis <- get_map(c(lon = centroid[2], lat = centroid[1]),
  zoom = 6, maptype = "terrain", color = "color", scale = "auto"
) %>%
  ggmap() +
  geom_sf(aes(fill = layer, alpha = layer),
    data = layer_poly_sf, inherit.aes = FALSE, lwd = 0
  ) +
  my_maptheme() +
  theme(panel.grid.major = element_line(color = "white")) +
  scale_fill_viridis_c("Grass\nPollen") +
  scale_alpha_continuous(range = c(0.1, 0.9)) +
  scale_x_continuous(limits = long_zoom, expand = c(0, 0)) +
  scale_y_continuous(limits = lat_zoom, expand = c(0, 0)) +
  guides(alpha = "none") +
    labs(
    x = "Lon",
    y = "Lat",
    title = "Grass Pollen in Switzerland",
    subtitle = "Hourly Average Concentration on the 1st of July 2020 at Midnight
    "
  )

ggmap_viridis
```

```{r eval = FALSE, include = FALSE}
ggsave(paste0(path_out, "ggmap_viridis.png"), ggmap_viridis, dpi = "retina")
```

## Plots for Swiss Municipalities

https://timogrossenbacher.ch/2019/04/bivariate-maps-with-ggplot2-and-sf/ Timo has a great blog about plotting maps in CH. The following open source shapefiles and code chunks are directly from his git repo here: https://github.com/grssnbchr/bivariate-maps-ggplot2-sf
Swisstopo has some great (free) map material for Switzerland which we will use here.

```{r echo = TRUE}
# read cantonal borders
canton_geo <- read_sf(paste0(path_in, "g2k15.shp"))
# read country borders - masking with read_sf didn't work
country_geo <- readOGR(paste0(path_in, "g2l15.shp"))
# read lakes
lake_geo <- read_sf(paste0(path_in, "g2s15.shp"))
# read productive area (2324 municipalities)
municipality_prod_geo <- read_sf(paste0(path_in, "gde-1-1-15.shp"))

# read in raster of relief
relief <- raster(paste0(path_in, "02-relief-ascii.asc")) %>%
  # hide relief outside of Switzerland by masking with country borders
  mask(country_geo) %>%
  as("SpatialPixelsDataFrame") %>%
  as.data.frame() %>%
  rename(value = `X02.relief.ascii`)

# clean up
rm(country_geo)
```

The mod_pollen data in this case was retrieved using fieldextra (MCH-Software). Which allows to extract modeled concentrations at any coordinate in the model domain.
The postprocessing of the fieldextra output is not yet streamlined. This could still be developed at some point.

```{r }
# Varnames from FieldExtra can be very long -> data.table
data <- data.table::fread(paste0(path_in, "c1_pollen_muni.csv")) %>%
  slice(1) %>%
  pivot_longer(ch0001:ch6810, names_to = "bfs_id", values_to = "mean") %>%
  select(bfs_id, mean) %>%
  mutate(bfs_id = as.integer(str_replace_all(bfs_id, "ch", "")))

municipality_prod_geo %<>%
  left_join(data, by = c("BFS_ID" = "bfs_id"))

# define number of classes
no_classes <- 6

# extract quantiles
quantiles <- municipality_prod_geo %>%
  pull(mean) %>%
  quantile(probs = seq(0, 1, length.out = no_classes + 1)) %>%
  as.vector() # to remove names of quantiles, so idx below is numeric

# here we create custom labels
labels <- imap_chr(quantiles, function(., idx) {
  return(paste0(
    round(quantiles[idx], 0),
    " - ",
    round(quantiles[idx + 1], 0),
    " m^-3"
  ))
})

# we need to remove the last label
# because that would be something like "- NA"
labels <- labels[seq_len(length(labels)) - 1]

# here we actually create a new
# variable on the dataset with the quantiles
municipality_prod_geo %<>%
  mutate(mean_quantiles = cut(mean,
    breaks = quantiles,
    labels = labels,
    include.lowest = T
  ))
```

```{r echo = TRUE}
map_timo <- ggplot(
  # define main data source
  data = municipality_prod_geo
) +
  # first: draw the relief
  geom_raster(
    data = relief,
    inherit.aes = FALSE,
    aes(
      x = x,
      y = y,
      alpha = value
    )
  ) +
  # use the "alpha hack" (as the "fill" aesthetic is already taken)
  scale_alpha(
    name = "",
    range = c(0.6, 0),
    guide = F
  ) + # suppress legend
  # add main fill aesthetic
  # use thin white stroke for municipality borders
  geom_sf(
    mapping = aes(
      fill = mean_quantiles
    ),
    color = "white",
    size = 0.1
  ) +
  # use the Viridis color scale
  scale_fill_viridis(
    option = "viridis",
    name = "Poaceae",
    alpha = 0.8, # make fill a bit brighter
    begin = 0.3, # this option seems to be new (compared to 2016):
    # with this we can truncate the
    # color scale, so that extreme colors (very dark and very bright) are not
    # used, which makes the map a bit more aesthetic
    end = 0.9,
    discrete = T, # discrete classes, thus guide_legend instead of _colorbar
    direction = 1, # dark is lowest, yellow is highest
    guide = guide_legend(
      keyheight = unit(5, units = "mm"),
      title.position = "top",
      reverse = T # display highest income on top
    )
  ) +
  # use thicker white stroke for cantonal borders
  geom_sf(
    data = canton_geo,
    fill = "transparent",
    color = "#ffffff",
    size = 0.5
  ) +
  # draw lakes in light blue
  geom_sf(
    data = lake_geo,
    fill = "#e4e6db6e",
    color = "transparent"
  ) +
  # add titles
  labs(
    x = NULL,
    y = NULL,
    title = "Grass Pollen in Switzerland",
    subtitle = paste("Hourly Average Concentration on",
       "the 1st of July 2020 at Midnight")
  ) +
  # add theme
  my_maptheme()
map_timo
```

```{r eval = FALSE, include = FALSE}
ggsave(paste0(path_out, "map_timo.png"), map_timo, dpi = "retina")
```
### Interactive

The following creates an interactive map (html widget) where the background can be chosen from a large variety of map providers.
If the polygon layer is fine or dense, plotting can take a long time or even crash. It's usually better to first aggregate the polygons.

```{r leaflet, echo = TRUE, fig.height = 5, fig.width = 7}
pal <- colorNumeric("YlOrRd", layer_poly_leaflet$layer)
mymap <- "https://stamen-tiles-{s}.a.ssl.fastly.net/terrain/{z}/{x}/{y}{r}.png"

leaflet_map <- leaflet(layer_poly_leaflet) %>%
  # More maps here: https://leaflet-extras.github.io/leaflet-providers/preview/
  addTiles(urlTemplate = mymap) %>%
  # Cosmo 1E Domain with extended boundaries
  fitBounds(lat1 = 42.60, lat2 = 49.60, lng1 = 0.10, lng2 = 16.80) %>%
  addPolygons(
    weight = 0,
    popup = as.character(round(values(layer_raster_vcoarse), 2)),
    smoothFactor = 0.5, fillColor = ~ pal(layer), fillOpacity = 0.6
  ) %>%
  addLegend(pal = pal, values = ~ layer_poly_leaflet$layer, title = "Poaceae")

leaflet_map
```

```{r eval = FALSE, include = FALSE}
saveWidget(leaflet_map,
  paste0(path_out, "leaflet_map.html"),
  selfcontained = TRUE
)
# This requires phantomjs for static map screenshot
mapshot(leaflet_map, file = paste0(path_out, "map_leaflet.png"))
```

And to conclude an interactive map with municipality level data.

```{r }
pal <- colorNumeric("YlOrRd", municipality_prod_geo$mean)

leaflet_map_muni <- leaflet(municipality_prod_geo %>%
  select(mean) %>%
  st_transform(crs = "+proj=longlat +datum=WGS84")) %>%
  # If the map is too heavy -> simplify shapes
  # ms_simplify(keep = 0.05))  %>%
  addTiles(urlTemplate = mymap) %>%
  # Cosmo 1E Domain with extended boundaries
  fitBounds(lat1 = 45, lat2 = 48, lng1 = 6, lng2 = 11) %>%
  addPolygons(
    weight = 0,
    popup = paste("<br/>",
      paste0("<b>", municipality_prod_geo$Secondary_, "</b><br/>"),
      paste0(as.character(round(municipality_prod_geo$mean, 0)),
        " Pollenkörner pro m³")
    ),
    smoothFactor = 0.5,
    fillColor = ~ pal(mean),
    fillOpacity = 0.6
  ) %>%
  addLegend(pal = pal, values = ~ municipality_prod_geo$mean, title = "Poaceae")

leaflet_map_muni
```

```{r eval = FALSE, include = FALSE}
saveWidget(leaflet_map_muni,
  paste0(path_out, "leaflet_map_muni.html"), selfcontained = TRUE)
```