---
title: "Plotting Maps in R"
author: "Simon Adamov"
date: "November 9, 2020"
output: html_document
---

This vignette shows some examples how to plot pretty maps in R with the latest packages.
In order to run the vignette on Tsa please `load module udunits` `module load gdal`, for the sf-package and `load module netcdf` for the ncdf4-package.
If the chunks are run outside of Rstudio in the terminal, then also install and load module pandoc.

User Inputs:

```{r}
# setwd("~/RProjects/pollenmaps")
path_in <- paste0(here::here(), "/input/")
path_out <- paste0(here::here(), "/output/")
nc_path <- "/users/paa/6_project/27_alnu_betu_kenda_tests/3_analysis_for_ESA/2_eps_quantile_differences/pollen_Q50_2020052006_0000.nc"

```

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE, 
                      message=FALSE,
                      fig.height=10, 
                      fig.width=16, 
                      fig.retina=3)

library(ncdf4)
library(dplyr)
library(broom)
library(leaflet)
library(raster)
library(OpenStreetMap)
library(ggplot2)
library(ggthemes)
library(htmlwidgets)
library(sf)

# remotes::install_version("caTools", version = "1.14")

```


```{r data}
con <- nc_open(nc_path)
# Print general info of file content
# summary(con)
# names(con$dim)
# names(con$var)

# get POAC
layer <- ncvar_get(con, "POAC")
x <- ncvar_get(con, "lon_1")
y <- ncvar_get(con, "lat_1")

nc_close(con)

r <- raster(layer, ymn = 42.67, ymx = 49.52, xmn = 0.16, xmx = 16.75, crs = "+proj=longlat")

layer_latlong <- tibble(x = c(x),
                       y = c(y),
                       layer = c(layer)) %>%
  mutate(x = if_else(x > 180, x - 360, x)) # if lons.dom refers to COSMO-1, lons range from 0 to 360 but lons west of 0 deg. need to be negative. hence make them negative here

#### GRID COSMO-1E ####
# This would load the Cosmo 1 Grid
# grid_c1e_path <- "/users/paa/11_resources/HSURF_RLON_RLAT_c1e.txt"
# gridboxes_lon <- 1170
# gridboxes_lat <- 786
# x <- scan(grid_c1e_path, skip = 919632,  n = gridboxes_lon * gridboxes_lat)
# y <- scan(grid_c1e_path, skip = 1839253, n = gridboxes_lon * gridboxes_lat)

```

```{r wrangling}

layer_raster <- rasterize(cbind(c(x), c(y)), r, layer_latlong$layer, fun = mean) # Convert the banana-shape into an equally spaced grid
layer_raster_coarse <- raster::aggregate(layer_raster, 25) # This is necessary otherwise the plots become way too heavy - adjust second agrument accordingly
layer_poly <- rasterToPolygons(layer_raster_coarse) # Polygons are usually faster to plot than raster images and avoid dependency on libpng
# layer_poly_coarse <- rmapshaper::ms_simplify(layer_poly) # Requires R-4.0.0

# These three lines prepare data for ggplot - workaround to avoid using the sf-package and its dependencies
# This works well with the opendata approach below and autoplot but takes a long time
layer_poly@data$id <- 1:nrow(layer_poly@data)
poly_fort <- broom::tidy(layer_poly, data = layer_poly@data)
poly_fort_mer <- merge(poly_fort, layer_poly@data, by.x = 'id', by.y = 'id')  # join data

# This is very efficient and works well with ggmap
layer_poly_sf <- st_as_sf(layer_poly)

```

```{r leaflet, eval = FALSE}

pal <- colorNumeric("YlOrRd", layer_latlong$layer)

leaflet_map <- leaflet(layer_poly) %>%
  addTiles(urlTemplate = 'https://stamen-tiles-{s}.a.ssl.fastly.net/terrain/{z}/{x}/{y}{r}.png') %>% # Check this for all background maps: https://leaflet-extras.github.io/leaflet-providers/preview/
  fitBounds(lat1 = 42.60, lat2 = 49.60, lng1 = 0.10, lng2 = 16.80) %>% # Cosmo 1E Domain with extended boundaries
  addPolygons(weight = 0, popup = as.character(round(values(layer_raster_coarse), 2)), smoothFactor = 0.5, fillColor = ~pal(layer), fillOpacity = 0.8) %>%
  # addRasterImage(layer_raster_coarse, opacity = 0.6) # Missing Libraries on Tsa - Reinstall packages with latest libpng module loaded
  # addCircles(lng = ~lon, lat = ~lat, col = ~layer, weight = 0, radius = 600, opacity = 0.3) # The circles approach crashes Rstudio - too many circles!
  # %>%
  addLegend(pal = pal, values = ~layer_latlong$layer, title = "Poaceae")

leaflet_map

# saveWidget(leaflet_map, paste0(path_out, "leaflet_map.html"), selfcontained = TRUE)
# This requires pandoc to be loaded on tsa - 1: module load EasyBuild-custom/cscs 2: eb Pandoc-2.10.eb –r 3: module load pandoc/2.10
# mapview::mapshot(leaflet_map, file = "/users/sadamov/map_leaflet.png") # Create png snapshot of map, requires R-3.6.0

```


```{r }
theme_map <- function(...) {
  theme_minimal() +
  theme(
    text = element_text(family = "Ubuntu Regular",
                        color = "#22211d"),
    # remove all axes
    axis.line = element_blank(),
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank(),
    # add a subtle grid
    panel.grid.major = element_line(color = "#dbdbd9", size = 0.2),
    panel.grid.minor = element_blank(),
    # background colors
    plot.background = element_rect(fill = "#f5f5f2",
                                   color = NA),
    panel.background = element_rect(fill = "#f5f5f2",
                                    color = NA),
    legend.background = element_rect(fill = "#f5f5f2",
                                     color = NA),
    # borders and margins
    plot.margin = unit(c(.5, .5, .2, .5), "cm"),
    panel.border = element_blank(),
    panel.spacing = unit(c(-.1, 0.2, .2, 0.2), "cm"),
    # titles
    legend.title = element_text(size = 11),
    legend.text = element_text(size = 9, hjust = 0,
                               color = "#22211d"),
    plot.title = element_text(size = 15, hjust = 0.5,
                              color = "#22211d"),
    plot.subtitle = element_text(size = 10, hjust = 0.5,
                                 color = "#22211d",
                                 margin = margin(b = -0.1,
                                                 t = -0.1,
                                                 l = 2,
                                                 unit = "cm"),
                                 debug = F),
    # captions
    plot.caption = element_text(size = 7,
                                hjust = .5,
                                margin = margin(t = 0.2,
                                                b = 0,
                                                unit = "cm"),
                                color = "#939184"),
    ...
  )
}
```


```{r efficient}

upperLeft  <- c(51, 0)
lowerRight <- c(40, 17.5)
map <- openmap(upperLeft, lowerRight, type = "stamen-terrain") # osm and other map providers available
map_proj <- openproj(map) # The raster map is in a mercator projection and must be transformed
raster_map <- autoplot(map_proj) +
  theme_map() +
  geom_polygon(data = poly_fort_mer, 
               aes(x = long, y = lat, group = group, fill = layer), 
               alpha = 0.5, 
               size = 0) +
  scale_fill_gradientn("Poac", colours = rev(heat.colors(200)))

raster_map
# ggsave(paste0(path_out, "raster_map.png"), raster_map, dpi = "retina") # Change resolution here

```


```{r}

# # Full Domain
# upperLeft_zoom  <- c(51, 0)
# lowerRight_zoom <- c(41, 18)

# Switzerland
upperLeft_zoom  <- c(48, 6)
lowerRight_zoom <- c(45.5, 10)

# # Zoom
# upperLeft_zoom  <- c(47.5, 8)
# lowerRight_zoom <- c(46, 10.5)
map_zoom <- openmap(upperLeft_zoom, lowerRight_zoom, type = "stamen-terrain") # osm and other map providers available
map_proj_zoom <- openproj(map_zoom) # The raster map is in a mercator projection and must be transformed

layer_latlong_zoom <- layer_latlong %>%
  filter(between(x, upperLeft_zoom[2], lowerRight_zoom[2]),
         between(y, lowerRight_zoom[1], upperLeft_zoom[1]))

alpha_low <- min(0.02/(lowerRight_zoom[2] - upperLeft_zoom[2]), 0.005)
alpha_high <- min(0.4/(lowerRight_zoom[2] - upperLeft_zoom[2]), 0.5)

gg_map <- autoplot(map_proj_zoom) +
  theme_map() +
  geom_point(aes(x = x, y = y, col = layer, alpha = layer), stroke = 1, shape = 15, size = 3.3, data = layer_latlong_zoom) +
  scale_alpha_continuous(range = c(alpha_low, alpha_high)) +
  scale_color_gradientn("Poac", colours = rev(heat.colors(200))) +
  guides(alpha = "none")

gg_map
# ggsave(paste0(path_out, "gg_map.png"), gg_map, dpi = "retina") # Change resolution here
```


The next example requires a Google account to retrieve data from the Google API. This package is quite powerful and requires very little pre

```{r ggmap, eval = TRUE}
# ?register_google
library(ggmap) # This package is more powerful but requires a Google account to query the Google API
register_google(key = "AIzaSyC-8Tyjbj1zc5fKdKZdbWaJk1H8CXdUIB8")

centroid  <- c(45, 8)
# Zoom
lat_zoom  <- c(44, 48)
long_zoom <- c(0, 10)

ggmap_heat <- get_map(c(lon = centroid[2], lat = centroid[1]), zoom = 6, maptype = "terrain", color = "color", scale = "auto") %>% 
  ggmap() +
  geom_sf(aes(fill = layer, alpha = layer), data = layer_poly_sf, inherit.aes = FALSE, lwd = 0) +
  theme_map() +
  theme(panel.grid.major = element_line(color = "white")) +
  scale_fill_gradientn("Grass\nPollen", colors = rev(heat.colors(200))) +
  scale_alpha_continuous(range = c(0.1, 0.9)) +
  scale_x_continuous(limits = long_zoom, expand = c(0, 0)) +
  scale_y_continuous(limits = lat_zoom, expand = c(0, 0)) +
  guides(alpha = "none")

ggmap_heat

# ggsave(paste0(path_out, "ggmap_heat.png"), ggmap_heat, dpi = "retina") # Change resolution here

```

Changing the color gradient is easy. https://www.datanovia.com/en/blog/top-r-color-palettes-to-know-for-great-data-visualization/

```{r}
centroid  <- c(45, 8)
# Zoom
lat_zoom  <- c(44, 48)
long_zoom <- c(0, 10)

ggmap_viridis <- get_map(c(lon = centroid[2], lat = centroid[1]), zoom = 6, maptype = "terrain", color = "color", scale = "auto") %>% 
  ggmap() +
  geom_sf(aes(fill = layer, alpha = layer), data = layer_poly_sf, inherit.aes = FALSE, lwd = 0) +
  theme_map() +
  theme(panel.grid.major = element_line(color = "white")) +
  scale_fill_viridis_c("Grass\nPollen") +
  scale_alpha_continuous(range = c(0.1, 0.9)) +
  scale_x_continuous(limits = long_zoom, expand = c(0, 0)) +
  scale_y_continuous(limits = lat_zoom, expand = c(0, 0)) +
  guides(alpha = "none")

ggmap_viridis

# ggsave(paste0(path_out, "ggmap_viridis.png"), ggmap_viridis, dpi = "retina") # Change resolution here
```

https://timogrossenbacher.ch/2019/04/bivariate-maps-with-ggplot2-and-sf/ Timo has a great blog about plotting maps in CH. The following open source shapefiles and code chunks are directly from his git repo here: https://github.com/grssnbchr/bivariate-maps-ggplot2-sf
Swisstopo has some great (free) map material for Switzerland which we will use here.

```{r }

library(lintr) # code linting
library(cowplot) # stack ggplots
library(viridis)
library(readr)
library(magrittr)
library(purrr)

# read cantonal borders
canton_geo <- read_sf(paste0(path_in, "g2k15.shp"))

# read country borders - masking with read_sf didn't work
country_geo <- rgdal::readOGR(paste0(path_in, "g2l15.shp"))

# read lakes
lake_geo <- read_sf(paste0(path_in, "g2s15.shp"))

# read productive area (2324 municipalities)
municipality_prod_geo <- read_sf(paste0(path_in, "gde-1-1-15.shp"))

# read in raster of relief
relief <- raster(paste0(path_in, "02-relief-ascii.asc")) %>%
  # hide relief outside of Switzerland by masking with country borders
  mask(country_geo) %>%
  as("SpatialPixelsDataFrame") %>%
  as.data.frame() %>%
  rename(value = `X02.relief.ascii`)

# clean up
rm(country_geo)

```

```{r }
data <- read_csv("input/data.csv")

municipality_prod_geo %<>%
  left_join(data, by = c("BFS_ID" = "bfs_id"))

# define number of classes
no_classes <- 6

# extract quantiles
quantiles <- municipality_prod_geo %>%
  pull(mean) %>%
  quantile(probs = seq(0, 1, length.out = no_classes + 1)) %>%
  as.vector() # to remove names of quantiles, so idx below is numeric

# here we create custom labels
labels <- imap_chr(quantiles, function(., idx){
  return(paste0(round(quantiles[idx] / 1000, 0),
                             "k",
                             " – ",
                             round(quantiles[idx + 1] / 1000, 0),
                             "k"))
})

# we need to remove the last label 
# because that would be something like "478k - NA"
labels <- labels[1:length(labels) - 1]

# here we actually create a new 
# variable on the dataset with the quantiles
municipality_prod_geo %<>%
  mutate(mean_quantiles = cut(mean,
                               breaks = quantiles,
                               labels = labels,
                               include.lowest = T))


```

```{r }

map_timo <- ggplot(
  # define main data source
  data = municipality_prod_geo
) +
  # first: draw the relief
  geom_raster(
    data = relief,
    inherit.aes = FALSE,
    aes(
      x = x,
      y = y,
      alpha = value
    )
  ) +
  # use the "alpha hack" (as the "fill" aesthetic is already taken)
  scale_alpha(name = "",
              range = c(0.6, 0),
              guide = F) + # suppress legend
  # add main fill aesthetic
  # use thin white stroke for municipality borders
  geom_sf(
    mapping = aes(
      fill = mean_quantiles
      ),
    color = "white",
    size = 0.1
  ) +
  # use the Viridis color scale
  scale_fill_viridis(
    option = "magma",
    name = "Average\nincome in CHF",
    alpha = 0.8, # make fill a bit brighter
    begin = 0.1, # this option seems to be new (compared to 2016):
    # with this we can truncate the
    # color scale, so that extreme colors (very dark and very bright) are not
    # used, which makes the map a bit more aesthetic
    end = 0.9,
    discrete = T, # discrete classes, thus guide_legend instead of _colorbar
    direction = 1, # dark is lowest, yellow is highest
    guide = guide_legend(
     keyheight = unit(5, units = "mm"),
     title.position = "top",
     reverse = T # display highest income on top
  )) +
  # use thicker white stroke for cantonal borders
  geom_sf(
    data = canton_geo,
    fill = "transparent",
    color = "white",
    size = 0.5
  ) +
  # draw lakes in light blue
  geom_sf(
    data = lake_geo,
    fill = "#D6F1FF",
    color = "transparent"
  ) +
  # add titles
  labs(x = NULL,
         y = NULL,
         title = "Switzerland's regional income",
         subtitle = "Average yearly income in Swiss municipalities, 2015") +
  # add theme
  theme_map()

map_timo

```